Toward Resolved Simulations of Burning Fronts in Thermonuclear X-ray Bursts

### 2018-09-16 astronum_2018/astronum-xrb.tex
some updates

\newcommand{\Uc}{{\,\bm{\mathcal{U}}}}
\newcommand{\Advs}[1]{\boldsymbol{\mathcal{A}} \left(#1\right)}

@@

\author{M. Zingale$^1$,
        K.~Eiden$^1$,
        Y.~Cavecchi$^2$,
        A.~Harpole$^1$,
        M.~P. Katz$^5$,
        J.~B. Bell$^2$,
        M.~Chang$^1$,
        C.~M. Malone$^6$, and
        W. Zhang$^2$}

@@

Describe resolution needs and flame structure, Figure~\ref{fig:xrb}

@@

As an example of the improvements in coupling between hydro and
reactions, Figure~\ref{fig:sdc} shows the mass fraction of helium
behind a detonation over the course of 3 timesteps.  The points
represent individual calls to the righthand side function of our
reaction network (a 13-isotope $\alpha$-chain), and the density of
points indicates how hard the integrator is working.  We use the same
tolerances for both the Strang-split method and the SDC method.  For
the SDC method, we are using a second-order accurate method.  We
predict a time-centered advection term, $\Advs{\Uc}^{n+1/2}$ using
standard unsplit Godunov methods, but explicitly include a reactive
source term in the tracing of the interface states.  We can then solve
the reactive system, using this advective term as a
piecewise-constant-in-time source
\begin{equation}
\frac{d{\Uc}}{dt} = -\Advs{\Uc}^{n+1/2} + {\bf R}(\Uc)
\end{equation}
This system can be integrated with standard ODE methods.  To achieve
second-order accuracy, we need to iterate, using the updated $\Uc$ to
create a reactive source included in the predictor for the advective
update, and then reintegrate the system.  The result of this iteration
is that the advection sees the effects of reactions over the timestep
and the reactions see the effects of advection.  For the zone tracked
in Figure~\ref{fig:sdc}, each SDC iteration called the righthand side
function of the network 5 times less than the Strang case.  
We are continuing to develop this new integration strategy, with full
fourth-order in space and time reacting hydrodynamics methods to
follow shortly.

@@

>>>>

\newcommand{\Uc}{{\,\bm{\mathcal{U}}}}
\newcommand{\Adv}[1]{{\left [\boldsymbol{\mathcal{A}} \left(#1\right)\right]}}
\newcommand{\Advt}[1]{{\left [\mathcal{\tilde{A}} \left(#1\right)\right]}}
\newcommand{\Advs}[1]{\boldsymbol{\mathcal{A}} \left(#1\right)}

@@

\author{M. Zingale$^1$,
        K.~Eiden$^1$,
        Y.~Cavecchi$^2$,
        A.~Harpole$^1$,
        J.~B. Bell$^2$,
        M.~Chang$^1$,
        M.~P. Katz$^5$,
        C.~M. Malone$^6$, 
        D.~E. Willcox$^2$, and
        W. Zhang$^2$}

@@

Describe resolution needs and flame structure, Figure~\ref{fig:xrb} \MarginPar{Mike will update this}

@@

As an example of the improvements in coupling between hydro and \MarginPar{Don can add some comments about startup costs?}
reactions, Figure~\ref{fig:sdc} shows the mass fraction of helium
behind a detonation over the course of 3 timesteps.  The points
represent individual calls to the righthand side function of our
reaction network (a 13-isotope $\alpha$-chain), and the density of
points indicates how hard the integrator is working.  We use the same
tolerances for both the Strang-split method and the SDC method.  For
the SDC method, we are using a second-order accurate method.  We
predict a time-centered advection term, $\Adv{\Uc}^{n+1/2}_i$ using
standard unsplit Godunov methods, but explicitly include a reactive
source term in the tracing of the interface states.  We can then solve
the reactive system, using this advective term as a
piecewise-constant-in-time source
\begin{equation}
\frac{d{\Uc_i}}{dt} = -\Adv{\Uc}^{n+1/2}_i + {\bf R}(\Uc_i)
\end{equation}
This system can be integrated with standard ODE methods.  To achieve
second-order accuracy, we need to iterate, using the updated $\Uc$ to
create a reactive source included in the predictor for the advective
update, and then reintegrate the system.  The result of this iteration
is that the advection sees the effects of reactions over the timestep
and the reactions see the effects of advection.  For the zone tracked
in Figure~\ref{fig:sdc}, each SDC iteration called the righthand side
function of the network 5 times less than the Strang case.  
We are continuing to develop this new integration strategy, with full
fourth-order in space and time reacting hydrodynamics methods to
follow shortly.

@@

\section{Summary}
